{\bf Landslide: A New Race-Finding Tool for 15-410}

### 2015-02-10 irb/hurdle.tex
Update sign-up sheet; write solicitation email; update lecture

	\\
	\checkbox{We pass the ``hurdle'' test suite listed in the P2 handout (on at least a vast majority of runs). This includes {\tt startle}, {\tt agility\_drill}, {\tt cyclone}, {\tt join\_specific\_test}, {\tt thr\_exit\_join}.}

>>>>

	\\
	\checkbox{We pass the ``hurdle'' test suite listed in the P2 handout (on at least a vast majority of runs). This includes {\tt startle}, {\tt agility\_drill}, {\tt cyclone}, {\tt join\_specific\_test}, {\tt thr\_exit\_join}. We acknowledge that {\bf a P2 that cannot pass these tests is incomplete, and Landslide will not work on it at all.}}

### 2015-02-10 landslide-lecture/2015/lecture.tex
update landslide lecture
Update sign-up sheet; write solicitation email; update lecture

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2014, December 3}

@@

\begin{document}
\renewcommand{\inserttotalframenumber}{34}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

\begin{frame}{Outline}
	\textbf{Theory: Seeing race conditions in a new way}
	\begin{itemize}
		\item Case study (example)
		\item Tabular execution traces
		\item The execution tree
	\end{itemize}
	{\bf Technique: Systematic testing}
	% Say: A way of enumerating all possible interleavings of a concurrent test case.
	\begin{itemize}
		\item Preemption points
		\item Challenges and feasibility
	\end{itemize}
	{\bf Tool: Landslide}
	\begin{itemize}
		\item How it works
		\item Coping with state space explosion
		\item Automatically choosing preemption points
		\item User study (that's you!)
			% !!! SAY !!! : If you are struggling, this is not for you, etc.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Case Study}
	\begin{center}
		\begin{tabular}{llll}
			\begin{tabular}{l}
				Consumer thread \\
				\hline
				\\
				\texttt{mutex\_lock(mx);} \\
				\\
				\texttt{if~(!work\_exists())} \\
				\texttt{~~~~cond\_wait(cvar,~mx);} \\
				\texttt{work~=~dequeue();} \\
				\\
				\texttt{mutex\_unlock(mx);} \\
				\texttt{access(work->data);} \\
			\end{tabular}
			& & &
			\begin{tabular}{l}
				Producer thread \\
				\hline
				\\
				\texttt{mutex\_lock(mx);} \\
				\\
				\texttt{enqueue(work);} \\
				\texttt{signal(cvar);} \\
				\\
				\texttt{mutex\_unlock(mx);} \\
				\\
				\\
			\end{tabular}
		\end{tabular}
		\linegap

@@

			\small \texttt{work~=~dequeue();~} & & \\
			\small \texttt{unlock(mx);} & & \\
			\small \texttt{access(work->data);}& & \\
			\hline
		\end{tabular}
	\end{center}
\end{frame}
\begin{frame}{Thread Interleavings (``good'' case, again)}
	\begin{center}
		\begin{tabular}{|l|l|l|}
			\hline
			\cellcolor{thread1} {\bf Thread 1} & \cellcolor{thread2} {\bf Thread 2} & \cellcolor{thread3} {\bf Thread 3} \\
			\hline
			\small \texttt{lock(mx);} & & \\
			\small \texttt{if~(!work\_exists())} & & \\
			\small \texttt{~~~~wait(cvar,~mx);} & & \\

@@

\begin{frame}{Testing}
	How can programmers be confident in their code's correctness?
	\begin{itemize}
		\item Unit tests
		\begin{itemize}
			\item good for basic functionality, bad for concurrency
		\end{itemize}
		\item Stress tests
		\begin{itemize}
			\item state of the art in 15-410
		\end{itemize}
		\item Theorem proving
		\begin{itemize}
			\item heavy burden on the programmers
		\end{itemize}
		\item Releasing to paying customers and worrying about correctness later
	\end{itemize}
	\linegap

@@

\begin{frame}{Systematic Testing - The Big Picture}
	\textbf{Goal: Force the system to execute every possible interleaving.}
	% Say: checking for crashes/memory errors/etc in each one.
	\begin{itemize}
		\item On 1st execution, schedule threads arbitrarily until program ends.
		\item At end of each branch, rewind system and restart test.
		\item Artificially add preemptions to test different thread interleavings.
		\item Intuitively: Generate many ``tabular execution traces''.
	\end{itemize}
	\pause
	\linegap

@@

	Systematic tests are {\em parameterized} by the set of PPs.
	\begin{itemize}
		\item If there are $n$ PPs and $k$ threads, state space size is $n^k$.
		\item Need to choose the set of PPs very carefully for test to be effective.
	\end{itemize}
\end{frame}

@@

	State of the art tools hard-code a fixed set of preemption points.
	\begin{itemize}
		\item E.g., ``all thread library API calls'' or ``all kernel mutex locks/unlocks''
		\item Depending on length of test, completion time is unpredictable.
		\item More often, a subset is better in terms of time/coverage. 
	\end{itemize}
	\pause
	\linegap
	Current systematic testing model is not user-friendly.
	\begin{itemize}
		\item Tool: ``I want to use these PPs, but can't predict completion time.''
		\item User: ``I have 16 CPUs and 24 hours to test my program.''
	\end{itemize}
	\linegap
	{\em Stress testing offers this -- can we offer it too?}
\end{frame}

\begin{frame}{``Iterative Deepening'' of Preemption Points}
	Goal: Run the best tests for a given CPU budget.
	\linegap

@@

@@

\section{Landslide for the People}

@@

	Risks \& Benefits
	\begin{itemize}
		\item Benefit: Landslide may help you find/fix bugs, improving your grade!
		\item Risk: Landslide may find no bugs and be a waste of your time.
	\end{itemize}
	% Say: Accordingly, although I hope you will use landslide enough to get good results for me, you should also responsibly allocate time between testing with landslide versus actually implementing your P2.
\end{frame}

% IRB approved, plan on publishing collected data, anonymized
% Expect to spend X time
% Participation has no impact on grade but you will find bugs

>>>>

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2015, February 11}

@@

\begin{document}
\renewcommand{\inserttotalframenumber}{36}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

\begin{frame}{Outline}
	\textbf{Theory: Seeing race conditions in a new way}
	\begin{itemize}
		\item Case study (example)
		\item Tabular execution traces
		\item The execution tree
	\end{itemize}
	{\bf Research Technique: ``Systematic testing''}
	% Say: A way of enumerating all possible interleavings of a concurrent test case.
	\begin{itemize}
		\item Preemption points
		\item Challenges and feasibility
	\end{itemize}
	{\bf Tool: Landslide}
	\begin{itemize}
		\item How it works
		\item Coping with state space explosion
		\item Automatically choosing preemption points
		\item User study (that's you!)
			% !!! SAY !!! : If you are struggling, this is not for you, etc.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Case Study}
	\begin{center}
		\begin{tabular}{llll}
			\fbox{
			\begin{tabular}{l}
				Consumer thread \\
				\hline
				\\
				\texttt{mutex\_lock(mx);} \\
				\\
				\texttt{if~(!work\_exists())} \\
				\texttt{~~~~cond\_wait(cvar,~mx);} \\
				\texttt{work~=~dequeue();} \\
				\\
				\texttt{mutex\_unlock(mx);} \\
				\texttt{access(work->data);} \\
			\end{tabular}
			}
			& & &
			\fbox{
			\begin{tabular}{l}
				Producer thread \\
				\hline
				\\
				\texttt{mutex\_lock(mx);} \\
				\\
				\texttt{enqueue(work);} \\
				\texttt{signal(cvar);} \\
				\\
				\texttt{mutex\_unlock(mx);} \\
				\\
				\\
			\end{tabular}
			}
		\end{tabular}
		\linegap

@@

			\small \texttt{work~=~dequeue();~} & & \\
			\small \texttt{unlock(mx);} & & \\
			\small \texttt{access(work->data);}& & \\
			\hline
		\end{tabular}
	\end{center}
\end{frame}
\begin{frame}{Thread Interleavings (different ``good'' case)}
	\begin{center}
		\begin{tabular}{|l|l|l|}
			\hline
			\cellcolor{thread1} {\bf Thread 1} & \cellcolor{thread2} {\bf Thread 2} & \cellcolor{thread3} {\bf Thread 3} \\
			\hline
			\small \texttt{lock(mx);} & & \\
			\small \texttt{if~(!work\_exists())} & & \\
			\small \texttt{~~~~wait(cvar,~mx);} & & \\

@@

\begin{frame}{Testing}
	How can programmers be confident in the correctness of their code?
	\begin{itemize}
		\item Unit tests
		\begin{itemize}
			\item good for basic functionality, bad for concurrency
		\end{itemize}
		\item Stress tests
		\begin{itemize}
			\item state of the art in 15-410
		\end{itemize}
		\item Theorem proving
		\begin{itemize}
			\item heavy burden on the programmers
		\end{itemize}
		\item Releasing to paying customers and worrying about correctness later
	\end{itemize}
	\linegap

@@

\begin{frame}{Systematic Testing - The Big Picture}
	\textbf{Goal: Force the system to execute every possible interleaving.}
	% Say: checking for crashes/memory errors/etc in each one.
	\begin{itemize}
		\item On 1st execution, schedule threads arbitrarily until program ends.
			\begin{itemize}
				\item This represents one branch of the tree.
			\end{itemize}
		\item At end of each branch, rewind system and restart test.
		\item Artificially add preemptions to produce different thread interleavings.
		\item Intuitively: Generate many ``tabular execution traces''.
	\end{itemize}
	\pause
	\linegap

@@

	Systematic tests are {\em parameterized} by the set of PPs.
	\begin{itemize}
		\item If there are $n$ PPs and $k$ threads, state space size is $n^k$.
		\item Need to choose the set of PPs very carefully for test to be effective.
		\begin{itemize}
			\item ``Effective'' = both comprehensive and feasible.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

	State of the art tools hard-code a fixed set of preemption points.
	\begin{itemize}
		\item E.g., ``all thread library API calls'' or ``all kernel mutex locks/unlocks''
		\item Depending on length of test, completion time is unpredictable.
		\item More often, a subset is better in terms of time/coverage. 
	\end{itemize}
	\pause
	\linegap
	Current systematic testing model is not user-friendly.
	\begin{itemize}
		\item Tool: ``I want to use these PPs, but can't predict completion time.''
		\item User: ``I have 16 CPUs and 24 hours to test my program.''
	\end{itemize}
	\linegap
	{\em Stress testing allows user to choose total run time -- can we offer this too?}
\end{frame}

\begin{frame}{Iterative Deepening of Preemption Points}
	Goal: Run the best tests for a given CPU budget.
	\begin{itemize}
		\item Technique: ``Iterative Deepening''
	\end{itemize}
	\linegap

@@

\section{Evaluation}

@@

\subsection{Landslide for the People}

@@

	Risks \& Benefits
	\begin{itemize}
		\item Benefit: Landslide may help you find/fix bugs, improving your grade!
		\item Risk: Landslide may find no bugs and be a waste of your time.
		\item Benefit: You might learn something...
	\end{itemize}
	% Say: Accordingly, although I hope you will use landslide enough to get good results for me, you should also responsibly allocate time between testing with landslide versus actually implementing your P2.
\end{frame}
